---
title: "Is There an Association Between Religion and Income"
date: "October 15 2014"
output:
  html_document:
    theme: cerulean
---

<!-- For more info on RMarkdown see http://rmarkdown.rstudio.com/ -->

<!-- Enter the code required to load your data in the space below. The data will be loaded but the line of code won't show up in your write up (echo=FALSE) in order to save space-->
```{r echo=FALSE}
library(ggplot2);library(plyr)
load(url("http://bit.ly/dasi_gss_data"))
```

<!-- In the remainder of the document, add R code chunks as needed -->

### Introduction:

Religious affilitation is still common in the United States. As we can see from the output below only 11% of Americans list None as their religion. This analysis will consider the interesting issue of religion and household income. We look at the following question. __Are religious affiliation and personal income are associated?__

```{r}
prop.table(table(gss$relig))*100
```


### Data:

Insert data section here...

### Exploratory data analysis:

Insert exploratory data analysis here...

### Inference:

Insert inference section here...

### Conclusion:

Insert conclusion here...

### References:
Smith, Tom W., Michael Hout, and Peter V. Marsden. General Social Survey, 1972-2012 [Cumulative File]. ICPSR34802-v1. Storrs, CT: Roper Center for Public Opinion Research, University of Connecticut /Ann Arbor, MI: Inter-university Consortium for Political and Social Research [distributors], 2013-09-11. doi:10.3886/ICPSR34802.v1
Persistent URL: http://doi.org/10.3886/ICPSR34802.v1